Topological Equivariant Artist Model for Visualization Library Architecture

### 2022-09-02 paper/main.tex
section 3.1

\subsubsection{Examples}
\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{tbundle.png}
  \caption{The mobius band is a non-trivial total space because the twist means that the fibers are sometimes [-1, 1] and other times [1,-1]. The mobius band can be represented as two trivial bundles, one with fiber [-1, 1] and one with fiber [1,-1] that can be glued on the edges into the band. The gluing rules ensure that the arrows on the edges go in the same direction.\label{fig:atct:nontrivialbundle}}
\end{figure}

The mobius band is an example of a non-trivial bundle because the twist in the middle, as shown in \autoref{fig:atct:nontrivialbundle}, means that there is at least one local .

But, as with all bundles, the mobius band can be decomposed into trivial bundle components, here the two sides of the mobius band. Each side has fibers going in the same direction, with the mismatch only apparent on the edges. These trivial bundles can be glued into the mobius band so long as the fibers (arrows on the edges) go in the same direction when glued. The data is encoded as the section of the bundle, here the sin curve.

\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{dbundle.png}
  \caption{The table from \autoref{fig:related-work:continuity:ktypes} has a 3 dimensional fiber (name, temperature, precipitation), a 0D base space, and the row is the section. The time series has a 2D fiber (time, precipitation) and a 1D base space since the data is sampled along a continuous space (physical time). Each section returns a data point along the line. The map is sparsely sampled points along a 2D plane encoding spatial continuity, and the fiber is a 3D space encoding (latitude, longitude, precipitation)
    \label{fig:atct:trivialbundle}}
\end{figure}

In \autoref{fig:atct:trivialbundle}, the base space $\dbase$ acts as an indexing space into the fiber space $\dfiber$. In the bundle encoding of the table, the indexing space is random keys; in the time series bundle, the base space is the interval [0,1]; and the map is sparse samples from a continuous space $[0,1]^{2}$ In contrast to a notion of a semantic binding between indexing space and field values, as proposed by Munzner\cite{munznerVisualizationAnalysisDesign2014}, the fields describing the continuity are part of the fiber. For example, the time is a fiber in the timeseries and the latitude and longitude are part of the map's fiber. This separation between connectivity and what it is called means the time or location can change units without a change to structure. It also provides a way to express data that may seem continuous but isn't, for example independent measurements over time. The data in \autoref{fig:atct:trivialbundle} comes from the same dataset; therefore we know that the bundles share connectivity and fiber space. It is expected that shared components be translated to visual elements in a consistent manner \cite{hullmanKeeping2018}, and in \autoref{sec:artist:operators} we introduce operators for expressing which components are shared and how to verify that they have been mapped into visual elements in a consistent manner.

\subsubsection{Uniform Abstract Graphic Representation}
One of the advantages of fiber bundles is that they are general enough that we can also encode the output of a visual algorithm as a bundle. This allows us to use the same structure to express the properties of data and the graphic that must be symmetric to the data in an equivariant (\autoref{sec:related-work:equivariance}) transformation. We denote the output as a graphic, but the use of bundles allows us to generalize to output on any display space, such as a screen or 3D print. 
\begin{equation}
  \label{eq:atct:fb:graphic}
  \begin{tikzcd}
      \gfiberc \arrow[r, hook, color=total] & \gtotalc \arrow[r, "\pi", two heads, color=total] & \gbasec
  \end{tikzcd}
\end{equation}
The total space \gtotalc\ is an abstraction of an ideal (infinite resolution) space into which the graphic can be rendered. The base space \gbasec\ is a parameterization of the display area, for example the inked bounding box in cairo \cite{CairographicsOrg}. The fiber space \gfiberc\ is an abstraction of the renderer fields, for example $\{x,\,y,\,r,\,g,\,b,\,a\}$.

As with data, we model the graphic generating functions as sections $\gsection$ of the graphic bundle 
\begin{equation}
  \label{eq:atct:fb_graphic_section}
  \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\}
\end{equation}
that map from a point in an openset in the graphic space $\gbasepoint \in \opensetg \in \gbase$ to a point in the graphic fiber $\gfiber$. The section evaluated on a single point $\gbasepoint$ returns a single graphic record, for example one pixel in an ideal resolution space. In our model, the unevaluated graphic section is passed to a renderer to generate graphics.

\begin{figure}[h!]
  \includegraphics[width=1\columnwidth]{fb_rho.png}
  \caption{For a 2D display a section $\gsection$ maps from each point $\gbasepoint$ into a fiber $\gfiber_{\gbasepoint}$ that encodes an RGB prerender space with infinite resolution $\mathcal{R}^{2}$. Each tiny colored box is an approximation of the return value of the same section function $\gsection$ evaluated on different points $\gbasepoint \in \gbase$ in the base space. We represent all $\gfiber_{\gbasepoint}$ as the fiber $\gfiber$ because a 2D bundle is a trivial bundle so $\gfiber_{\gbasepoint} = \gfiber$ for all \gbasepoint.
  \label{fig:atct:fb:graphic}}
\end{figure}

In \autoref{fig:atct:fb:graphic}, the section function $\gsection$ maps into the fiber for a simplified 2D RGB infinite resolution prerender space and returns the $\{x,y,r,g,b\}$ values of a pixel in an infinite resolution space. In \autoref{fig:atct:fb:graphic} these pixels are approximated as the small orange and black colored boxes. Each box is the output of the $\gsection(\gbasepoint)$ section that intersects the box. The set of all pixels returned by a section evaluated on a given visual base space $\gsection|_{\gbase}$ can yield a visual element, such as a marker, line, or piece of a glyph. \autoref{fig:atct:fb:graphic} illustrates a highly idealized space with no overlaps. To support overlaps, we can add another fiber element $\gfiber_{z}$ and leave it to the renderer to appropriately blend layers.

>>>>

\subsubsection{Examples}
\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{trivnon.png}
  \caption{One local trivialization of both a trivial cylinder bundle and a non-trivial mobius bundle is a plane over the interval $\left[0,2\pi\right]$. The transition map that transforms this trivialization into a cylinder is an identity map aligning two \textit{up} fibers, while the transition map that transforms this trivialization into a mobius bundle aligns an \textit{up} fiber with a \textit{down} fiber.  
  \note{not sure about wording here/ maybe should still have two transition maps so non-trivial has 1 id and 1 not-id}
  \label{fig:atct:nontrivialbundle}}
\end{figure}

As shown in \autoref{fig:atct:nontrivialbundle}, a cylinder bundle and mobius band bundle over the same base space (here a unit circle) can be locally trivialized in the same manner. The local trivialization in \autoref{fig:atct:nontrivialbundle} is a plane over the interval $\left[0,2\pi\right]$. In the cylinder bundle, all fibers go in the same direction, \textit{up}, such that transition maps on local trivializations are all identity maps aliging two \textit{up} fibers. This property of transition maps is what charaterizes it as a trivial bundle, while local trivializations of  non-trivial bundles always have at least on non-identity transition map. In \autoref{fig:atct:nontrivialbundle} this is the sole transtion map and it aligns an \textit{up} fiber with a \textit{down} fiber, thereby creating the twist in the mobius band. The data is encoded as the section of the bundle, here exemplified as the sin curve. In the cylinder bundle, the sin curve is continuous on the whole space, while the sin curve is discontinuous at the twist in the mobius band and can be made continuous via the transition map.

\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{dbundle.png}
  \caption{The table from \autoref{fig:related-work:continuity:ktypes} has a 3 dimensional fiber (name, temperature, precipitation), a 0D base space, and each row is a section. Each time series is a section of a bundle with a 2D fiber (time, precipitation) and a 1D base spac encoding temporal continuity. Each location of the rain map is a section of a bundle with a 2D plane base encoding spatial continuity and a 3D fiber space encoding (latitude, longitude, precipitation)
    \label{fig:atct:trivialbundle}}
\end{figure}

In \autoref{fig:atct:trivialbundle}, the base space $\dbase$ acts as an indexing space into the fiber space $\dfiber$. In the bundle encoding of the table, the indexing space is arbitrarily numbered keys; in the time series bundle, the base space is the interval [0,1]; and the map is sparse samples from a continuous space $[0,1]^{2}$ In contrast to a notion of a semantic binding between indexing space and field values, as proposed by Munzner\cite{munznerVisualizationAnalysisDesign2014}, the fields describing the continuity are part of the fiber. For example, the time is a fiber in the timeseries and the latitude and longitude are part of the map's fiber. This separation between connectivity and what it is called means the time or location can change units without a change to structure. It also provides a way to express data that may seem continuous but isn't, for example independent measurements over time. The data in \autoref{fig:atct:trivialbundle} comes from the same dataset; therefore we know that the bundles share connectivity and fiber space. It is expected that shared components be translated to visual elements in a consistent manner \cite{hullmanKeeping2018}, and in \autoref{sec:artist:operators} we introduce operators for expressing which components are shared and how to verify that they have been mapped into visual elements in a consistent manner.

\subsubsection{Uniform Abstract Graphic Representation}
One of the advantages of fiber bundles is that they are general enough that we can also encode the output of a visual algorithm as a bundle. This allows us to use the same structure to express the properties of data and the graphic that must be symmetric to the data in an equivariant (\autoref{sec:related-work:equivariance}) transformation. We denote the output as a graphic, but the use of bundles allows us to generalize to output on any display space, such as a screen or 3D print. 
\begin{equation}
  \label{eq:atct:fb:graphic}
  \begin{tikzcd}
      \gfiberc \arrow[r, hook, color=total] & \gtotalc \arrow[r, "\pi", two heads, color=total] & \gbasec
  \end{tikzcd}
\end{equation}
The total space \gtotalc\ is an abstraction of an ideal (infinite resolution) space into which the graphic can be rendered. The base space \gbasec\ is a parameterization of the display area, for example the inked bounding box in cairo \cite{CairographicsOrg}. The fiber space \gfiberc\ is an abstraction of the renderer fields; for example a 2 dimension screen has pixels that can be parameterized $ \gfiber=\{x,\,y\,z\,r,\,g,\,b,\,a\}$.

As with data, we model the graphic generating functions as sections $\gsection$ of the graphic bundle 
\begin{equation}
  \label{eq:atct:fb_graphic_section}
  \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\}
\end{equation}
that map from a point in an openset in the graphic space $\gbasepoint \in \opensetg \subseteq \gbase$ to a point in the graphic fiber $\gfiber$. The section evaluated on a single point $\gbasepoint$ returns a single graphic record, for example one pixel in an ideal resolution space. In our model, the unevaluated graphic section is passed to a renderer to generate graphics.

\begin{figure}[h!]
  \includegraphics[width=1\columnwidth]{fb_rho.png}
  \caption{For a 2D display, a section $\gsection$ maps from each point $\gbasepoint$ into a fiber $\gfiber$ that encodes an RGB prerender space with infinite resolution $\mathcal{R}^{2}$. Each tiny colored box is an approximation of the return value of the same section function $\gsection$ evaluated on different points $\gbasepoint \in \gbase$ in the base space. \note{add alpha and z channels} 
  \label{fig:atct:fb:graphic}}
\end{figure}

In \autoref{fig:atct:fb:graphic}, the section function $\gsection$ maps into the fiber for a simplified 2D RGB infinite resolution prerender space and returns the $\{x,y,r,g,b\}$ values of a pixel in an infinite resolution space. In \autoref{fig:atct:fb:graphic} these pixels are approximated as the small orange and black colored boxes. Each pixel is the output of the $\gsection(\gbasepoint)$ section that intersects the box. The set of all pixels returned by a section evaluated on a given visual base space $\gsection|_{\gbase}$ can yield a visual element, such as a marker, line, or piece of a glyph. \autoref{fig:atct:fb:graphic} illustrates a highly idealized space with no overlaps. Overlaps can be indicated via fiber element $\gfiber_{z}$ and it can be left to the renderer to appropriately blend layers based on $\gfiber_{z}$ and $\gfiber_{a}$.
